---
title: "daily"
author: "Hannah Reich"
date: "6/21/2021"
output: html_document
---

# 20210621, SCRIPT adapted FROM Jill https://github.com/JillAshey/Astrangia_repo/blob/main/scripts/Daily.Rmd 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
library(tidyverse)
library(gridExtra)
```

# Check working directory
```{r}
 setwd("C:/Users/roxyr/Documents/GitHub/Lab_Management/Lab_Resourses/CBLS_Aquarium_WQ_Tracking")
```
# Load data
``{r}
# Load data
daily <- read.csv("/Users/roxyr/Documents/GitHub/Lab_Management/Lab_Resourses/CBLS_Aquarium_WQ_Tracking/CBLS_watertables_waterquality_tracking.csv", header = T)
#daily$Timepoint <- as.Date(daily$Timepoint, "%m/%d/%Y")
daily$Tank_ID <- as.factor(daily$Tank_ID)
daily$Treatment_ID <- as.factor(daily$Treatment_ID)
```

# Discrete pH calculations from Tris calibrations.
```{r}
path <-("/Users/roxyr/Documents/GitHub/Lab_Management/Lab_Resourses/CBLS_Aquarium_WQ_Tracking/tris_cal") #set path to calibration files
file.names<-list.files(path = path, pattern = "csv$") #list all the file names in the folder to get only get the csv files
pH.cals <- data.frame(matrix(NA, nrow=length(file.names), ncol=3, dimnames=list(file.names,c("Date", "Intercept", "Slope")))) #generate a 3 column dataframe with specific column names
for(i in 1:length(file.names)) { # for every file in list start at the first and run this following function
  Calib.Data <-read.table(file.path(path,file.names[i]), header=TRUE, sep=",", na.string="NA", as.is=TRUE) #reads in the data files
  file.names[i]
  model <-lm(mVTris ~ TTris, data=Calib.Data) #runs a linear regression of mV as a function of temperature
  coe <- coef(model) #extracts the coeffecients
  summary(model)$r.squared #extracts the r squared
  plot(Calib.Data$mVTris, Calib.Data$TTris) #plots the regression data
  pH.cals[i,2:3] <- coe #inserts coefficients in the dataframe
  pH.cals[i,1] <- substr(file.names[i],1,8) #stores the file name in the Date column
}
colnames(pH.cals) <- c("Calib.Date",  "Intercept",  "Slope") #rename columns
pH.cals #view data
#constants for use in pH calculation 
R <- 8.31447215 #gas constant in J mol-1 K-1 
F <- 96485.339924 #Faraday constant in coulombs mol-1
#merge with Seawater chemistry file
SW.chem <- merge(daily, pH.cals, by="Calib.Date")
```

# Calculate total pH. 
```{r}
mvTris <- SW.chem$Temperature*SW.chem$Slope+SW.chem$Intercept #calculate the mV of the tris standard using the temperature mv relationships in the measured standard curves 
STris<-35 #salinity of the Tris
phTris<- (11911.08-18.2499*STris-0.039336*STris^2)*(1/(SW.chem$Temperature+273.15))-366.27059+ 0.53993607*STris+0.00016329*STris^2+(64.52243-0.084041*STris)*log(SW.chem$Temperature+273.15)-0.11149858*(SW.chem$Temperature+273.15) #calculate the pH of the tris (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)
SW.chem$pH.Total<-phTris+(mvTris/1000-SW.chem$pH.MV/1000)/(R*(SW.chem$Temperature+273.15)*log(10)/F) #calculate the pH on the total scale (Dickson A. G., Sabine C. L. and Christian J. R., SOP 6a)

# SAVE SW CHEM FILE (HAS TOTAL PH)
```



